Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings

### 2019-08-28 Manuscript/manuscript.Rmd
edits

After obtaining our optimally built random forest classifier, we investigated whether our predictive models can alleviate resolution limitations of Hi-C data by predicting TAD boundaries at base pair resolution. We were able to identify regions where the model predicted the base pair to be a TAD boundary with probability equal to 1. [??? Where is this figure? It seemed important !!!how best to include? as an example for specific section of genome?] By clustering the regions (see Methods), we were able to predict a comparable number of TAD boundary regions as called by the ARROWHEAD TAD caller. However, only 24% and 29% of predicted and called (and 10 kb flanked) TAD boundaries overlapped for the GM12878 and K562 cell line, respectively, suggesting differing locations in TAD boundaries throughout the linear genome (Supplementary Figure 9). [??? Need further drill down, as we discussed; !!!discuss in meeting]

For each predicted TAD boundary region, we were able to identify the summit locations of TAD boundary location by performing partitioning around medoids (see Methods). To compare the summit locations of predicted and called TAD boundaries, we evaluated the signal strength of ChIP-seq peak regions around a 10 kb flanking region. We found that the signal strength for the CTCF, RAD21, SMC3, and ZNF143 [??? Make all in this order!; !!!reasoning of order is to match heatmap; discuss in meeting] TFBS peaked at our predicted TAD boudaries, while was poorly defined around the called TAD boundaries (Figure 5; Supplementary Figure 10).  Thus, our model is able to alleviate resolution limitations and precisely locate TAD boundaries that correspond to the highest signal strength of known TAD boundary-forming TFBSs.

@@

>>>>

After obtaining our optimally built random forest classifier, we investigated whether our predictive models can alleviate resolution limitations of Hi-C data by predicting TAD boundaries at base pair resolution. We were able to identify regions where the model predicted the base pair to be a TAD boundary with probability equal to 1. [??? Where is this figure? It seemed important !!!how best to include? as an example for specific section of genome?] By clustering the regions (see Methods), we were able to predict a comparable number of TAD boundary regions as called by the ARROWHEAD TAD caller. However, only 24% and 29% of predicted and called (and 10 kb flanked) TAD boundaries overlapped for the GM12878 and K562 cell line, respectively, suggesting differing locations in TAD boundaries throughout the linear genome (Supplementary Figure 9). [??? Need further drill down, as we discussed; !!!added distance to annotations figure in next paragraph]

For each predicted TAD boundary region, we were able to identify the summit locations of TAD boundary location by performing partitioning around medoids (see Methods). To compare the summit locations of predicted and called TAD boundaries, we evaluated the signal strength of ChIP-seq peak regions around a 10 kb flanking region. We found that the signal strength for the CTCF, RAD21, SMC3, and ZNF143 [??? Make all in this order!; !!!reasoning of order is to match heatmap; discuss in meeting] TFBS peaked at our predicted TAD boudaries, while was poorly defined around the called TAD boundaries (Figure 5; Supplementary Figure 10).  Thus, our model is able to alleviate resolution limitations and precisely locate TAD boundaries that correspond to the highest signal strength of known TAD boundary-forming TFBSs. Furthermore, when compared to called TAD boundaries, the boundaries that we predicted with our model were found to be statistically significantly closer to the top ranked TFBS, suggesting stronger enrichment profiles near predicted TAD boundaries (Supplementary Figure 11).

@@

**Supplementary Figure 11. Known architectural transcription factors were statistically significantly closer to predicted TAD boundaries then called TAD boundaries.** Barplots illustrate the log2 distance in basepairs from classes of TAD boundaries to transcription factors including CTCF, RAD21, SMC3, and ZNF143 for the GM12878 (A) and K562 (B) cell lines. TAD boundaries were classified into the 3 classes. Predicted but not called (PBNC) were the set of 10 kb flanked predicted TAD boundaries the did not overlap with 10 kb flanked called TAD boundaries. Called and predicted (CAP) were the set of 10 kb flanked called and predicted TAD boundaries that overlapped with each other. Called but not predicted (CBNP) were the set of 10 kb flanked called TAD boundaries that did not overlap with 10 kb flanked predicted TAD boundaries.
